TBA

### 2020-08-25 manuscript.tex
Andy pull

Over the last couple decades, DNA has proved to be much more than a vessel for genetic information. From sensing, to computing, to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen applications \citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems.pdf, Rothemund2006FoldingPatterns, Gu2010ALine}. Recently, structural DNA nanotechnology has enabled self-assembly on micro to milli scales, and dynamic DNA nanotechnology has been used to perform basic calculation and to probe single molecules via temporal DNA signatures\citep{MhatreV.HoJi-AnnLee2012NIHAccess, Bui2018, Shah2019}. \noter{Please correct reference ``MhatreV.HoJi-AnnLee2012NIHAccess''} Both technologies rely on the hybridization reaction between complementary DNA strands and leverage the flexibility of shorter DNA oligomers to participate in these reactions. Although many experimental and computational studies have rigorously explored DNA dynamical phenomena such as hybridization, hairpin formation, and single base pair flipping, the sequence-dependent mechanisms of hybridization and dissociation dynamics are not fully understood \citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition, Porschke1973ThermodynamicsPairs, Chen2007InfluenceHybridization, Craig1971ElaxationOligon}. Moreover, it is unclear the extent to which these processes evolve in an "all-or-nothing" fashion or if an ensemble meta-stable states facilitates the transition. Recent breakthrough studies have coupled experimental techniques with machine learning and MD simulations to investigate and predict sequence-dependent kinetics \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence}. Where these studies focus on association and dissociation kinetics alone, we broaden our analysis into higher order dynamical processes and meta-stable intermediates. The stability of certain intermediates states, such as out-of-register or  base pairing in repetitive sequences, has been well documented in previous computational studies \citep{Phys2014, Xiao2019}). Furthermore, frayed structures and dynamics have been investigated in numerous computational and experimental studies \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix}. Sandstead et al. highlighted the role of these dynamics during duplex dissociation, where the stability of frayed states was dictated by G:C bond placement in the 10-mer oligonucleotide sequence \citep{Sanstead2016}. In this work, we study the same four sequences explored by Sandstead et al in an effort to uncover the sequence-dependent dynamics and their relation to  metastable structures mentioned above.

\textcolor{purple}{I would suggest re-organizing this paragraph to emphasize how the dynamic insight from various experiments is limited and therefore high-level simulations are required to understand DNA dynamics in detail. Then, you can summarize why simulating DNA hybridization is difficult and the previously employed techniques} Given the long timescales on which DNA hybridization and dissociation events occur, the study of these processes are generally not amenable to direct simulation techniques \citep{Phys2014}. Instead, many previous studies of DNA hybridization have employed accelerated sampling methods such as umbrella sampling \citep{Schmitt2013ExploringSurface} transition path sampling \citep{Sambriski2009}, and forward flux sampling \citep{Schmitt2013ExploringSurface, Sambriski2009,  Hoefert2011MolecularOligonucleotides, Romano2013DNADependence}. Other computational studies use dramatically elevated temperature or denaturing solvent concentrations induce one-way dissociation events \citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding}. Experimental studies often employ temperature-jump, salt-jump, pH-jump, or other perturbative methods to drive DNA out of equilibrium and monitor relaxation processes in one direction \citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971ElaxationOligon, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization}. Single molecule diffusion and tethered multifluorophore assays facilitate equilibrium analysis, however these present technical difficulties and are hampered by data collection rates \citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices}. Förster resonance energy transfer (FRET) analysis, particularly when coupled with methods mentioned above, provides additional resolution, but it unclear how fluorescent tags may interfere with the dynamics of short nucleotides \citep{Morrison1993SensitiveSolution}.

@@

We initialized four sequences previously investigated by Sandstead et al. -- 5'ATATATATAT3' (AT-all), 5'GATATATATC3' (GC-end), 5'ATATGCATAT3' (GC-core), and 5'ATGATATCAT3' (GC-mix) -- along with their complementary strands according to 3SPN.2 documentation \citep{Sanstead2016, Phys2014}. We initialized explicit ions such that 240 mM NaCl and 18 mM MgCl2 \textcolor{purple}{So these are initially in the bulk and then may interact with the DNA? Are these the effective concentrations after you shrink the box?} were added to the box in addition to 18 Na counter ions to balance the charge from the 9 phosphate groups in each oligonucleotide backbone \citep{Hinckley2015}. In order to maximize concentration without allowing strands to see each other through periodic boundaries, we set the box size just larger than the sum of the maximum end-to-end extension length of a single strand and the force cutoff (using Ewald summation method is set at 20 A). This translated to a box size of 77.74 Å and an effective oligo concentration of 7 mM. We used an Ewald potential to calculate long range Coulombic interaction between DNA and ions. We maintained a Debye-Huckel screening potential with an ionic strength of 240 mM -- corresponding to NaCl concentration -- to account for phosphate backbone interactions. These electrostatics preserve the persistence length and intrinsic curvature of DNA while taking the effects of ion-DNA interactions\citep{Hinckley2015}. We run our simulations in the NVE ensemble and fix temperature via a Langevin thermostat in order to model implicit solvent interactions \citep{Schneider1978Molecular-dynamicsTransitions}. Simulation temperature was determined by sequence specific melting temperature such that strands were equally likely to spend time in the hybridized or dissociated state. Tmelt simulations were run for $3*10^{8}$ time steps, equating to 4.5$\mu$s simulation time, and frames were saved every 30 ps. A 15 fs timestep was used for all simulations runs. For each sequences, 100 simulations were performed in parallel, consuming about 32 serial cpu-hours of computation time for each independent simulation. An approximate Tmelt Boltzman distribution was replicated by initializing half of runs from the hybridized state and half from a random dissociated state. In order to allow for further equilibration, the first third (1.5 $\mu$s) of each simulation was removed  \textcolor{purple}{Why 1.5 microseconds? Is this arbitrary or is there a good reason to remove 1.5?}, resulting in 100 x 100000 frames and a total of 300$\mu$s simulation time per sequence.

@@

All intermolecular pairwise distances for both oligonucleotides were calculated at each frame using the MDtraj software package \citep{McGibbon2015MDTraj:Trajectories}. Based on the self-complementary nature of each sequence, we averaged permutable distances (45 pairs in total) together. This permutation reduction follows a similar procedure used in TICAgg coordinates construction \citep{Sengupta2019AutomatedSelf-assembly}. The VAMP-2 scoring method \textcolor{purple}{Since the VAMP-2 scoring comes up a lot, can you maybe another sentence or two on how exactly this works? Citing the Noe paper helps, but I think experimentalists may get a bit lost with it}, which represents the sum of squared estimates for the transfer operator, was employed to evaluate the quality of this feature set \citep{Mardt2018VAMPnetsKinetics}. We compared the VAMP-2 score for the permutation-free distances to the complete set of intermolecular distances and found only a small loss in kinetic variance when using the reduced data set. Furthermore, we found an increase in VAMP-2 score when using reciprocal pairwise distances and chose these reciprocal permutation-free coordinates as a consistent feature set. These features were normalized and passed into sequence-specific SRVs.

@@

\section{\label{sec:Results}Results}
- lead with same kind of diagram for each sequence, add additional figures when necessary for a given sequence

\subsection{\label{sec:Results}AT-all}

\subsubsection{\label{sec:Results}SRV optimization and analysis}

In our analysis, we found that the AT-all sequence, given its repetitive structure and lack of GC-content, produced the cleanest dynamics and displayed a canonical spectral gap between modes. For this reason, we lead our discussion with this sequence and use it as a case study to work through our SRV-MSM pipeline step-by-step. Our first task was to identify the SRV lag time that was longer than the intrinsic Markov timescales of the system, yet short enough to resolve the dynamics of interest \citep{Phys2018MarkovValidation}. We found that most implied timescales (with the exception of the leading timescale) converge at a lag time of 600 ns. We kept a looser constraint on the convergence of the leading mode as it reflects the less Markovian dynamic of strand association and dissociation \textcolor{purple}{Not sure how to describe this or if it warrants further discussion?} . It should be noted that our lag time selection was informed by the other sequences as well to maintain consistency. Next we selected an optimal number of SRV components to include in our analysis. After a certain point, higher order dynamical modes provide diminishing contributions the overall kinetic variance as measured by the Vamp-2 score, and the model can begin fitting on statistical noise in the trajectory data instead of the true dynamics \citep{McGibbon2015VariationalKinetics}. It is also more difficult to perform kmeans clustering on a high dimensional space, especially when those higher dimensions are less kinetically relevant \citep{Pande2010EverythingAsk}. For these reasons, the number of slow SRV components should be carefully selected based on the specific system of interest. As shown in (Figure \ref{fig:srv_crossval}), we see a clear convergence of the VAMP-2 score after five slow modes and select these modes as our optimized SRV basis. Next, we seek to interpret the physical relevance of these leading modes by plotting the Pearson correlation of each mode with the 100 intermolecular distances between strands. The quantitative meaning of these coordinates can be difficult to interpret given their nonlinear relationship to the SRV collective variables, but the relative different between these correlations shows which coordinates are most affected by (or effective on) each process. For example, the first slow mode shows a significant positive correlation to each distance and the strongest correlation with matching base pair distances (shown along the main diagonal). Given these relationships and the substantially longer timescale of this process, we can deduce that this leading mode corresponds to the dynamics of the overall hybridization and dissociation process. The next four SRV components all show a relatively high correlation along different offset diagonals. These diagonals correspond to the intermolecular distances between shifted base pairs and point to the existence of gradually faster shifting or "slithering" processes \textcolor{purple}{I suppose the GCGCGC sequence in the Wales paper provides an interesting comparison to AT-all
}. Slow slithering mechanisms have similarly been reported in simulation studies to occur on orders of magnitude longer timescales than underlying fast dynamics \citep{Markegard2015, Xiao2019, Maciejczyk2014DNAModel}.

\subsubsection{\label{sec:Results}SRV-MSM construction and optimization}

From this analysis we can determine that the slowest dynamics are fully characterized by the hybridization process and slithering behavior of out of register base pairs. Although this is qualitatively informative, we can access a more holistic picture of sequence kinetics and thermodynamics by using these SRV coordinates as a basis on which to construct an MSM. Because these coordinates are already capturing a majority of the system's kinetic variance, they serve as an ideal basis on which to group frames into microstates. We performed k-means clustering, and optimized the number of microstates at 200 by monitoring VAMP-2 score. Next, we selected an MSM lag time in a similar fashion to our SRV lag time selection process. In Figure \ref{fig:AT-all_dynamic} we compare the convergence of SRV-MSM and TICA-MSM timescales, where SRV-MSM timescales converge consistently faster and to higher values of all leading modes. This enables us to select a shorter lag time and build a higher resolution model than we could from an analogous TICA basis. Using an MSM lag time of 1.2 ns, we then built a Bayesian MSM to calculate transition probability matrix between each microstate. Finally, PCCA+ spectral clustering was implemented to group these microstates into macrostates that each represent a collection of metastable structures. Previous work have used a common set of microstates and/or performed manual clustering of microstates based on physical read outs from simulation data (stacking score, energies, etc) \citep{Pinamonti2017PredictingModels}, \citep{PinamontiAnalyzedModels}. Although these techniques are useful for performing comparisons between sequences, we saw better results when optimizing MSMs to capture the most detail of sequence individually and thus developed an independent set of microstates and macrostates for each sequence. For AT-all, we kept to the convention of clustering into n+1 macrostates, where n is the number of slow components captured by the MSMs. To visualize these six macrostates, we project the data into the two leading TICA coordinates. Although SRVs outperform these coordinates for the purpose of MSM construction, TICA represent good high variance collective variables on which to visualize free energies and state assignments \citep{Sidky} \textcolor{purple}{Still not sure why you use tICA here, does not nonlinear nature make them harder to include physically? Should these be in SI}. It is clear in Figure \ref{fig:AT-all_in_one} that the PCCA state assignments are capturing free energy minima in TICA space as independent states. After assigning these macrostates we can then calculate their relative probabilities and free energies, visualize representative molecular renderings, and estimates transition probabilities between states. In this final step, we use a minimum flux cutoff of 2e-6 in order to mitigate erroneous quick transition or skipping between states.

@@

\subsubsection{\label{sec:Results}Implied timescales for TICA-MSMs for }

In our coarse-grained SRV-MSM, we observe an "aligned hybridized" state, dissociated state, and four "shifted" states characterized by different combinations of out-of-register base pairings. These shifted states consist of 2 or 4 base pair shifting (single or doubled shifted) in either the 5' or 3' direction with varied stability and transition fluxes between kinetically neighboring states. This state decomposition is expected given the repetitive nature of the AT-all sequence and previous computational results that find shifted conformations form "deep kinetic traps" \citep{Xiao2019}, \citep{Phys2014}. We found a substantial difference in thermodynamic stability and kinetic behavior between the 5'A shifted states and 3'T shifted states. This difference can be accounted for by examining experimental studies on the thermodynamics of "dangling ends" -- unpaired bases adjacent to the paired duplex -- and "inert tails" -- free bases that extend beyond the dangling end \citep{Michele2014EHybridization}. Dangling 5' ends have a consistently stabilizing effect, and inert tails decrease stability as they increase in length \textcolor{purple}{The effect of inert tails seems sequence dependent and not always destabilizing (Benight), especially at high ionic strength}. It has been shown that two base pair 5' dangling ends (or dangling ends with one inert tail) have higher melting temperature and are enthalpically favorable compared to 3' ends \citep{Senior1988InfluenceDuplexes, Dickman2012ThermodynamicDNAs} \textcolor{purple}{The Benight paper also shows that 3'-dangling ends and tails alter the change in heat capacity (dCp) upon hybridization much more than 5'. Additionally, their analysis suggests 3'-dangling ends lead to a reduction in bound counterions relative to 5'. Taken together, it would seem that 3'-dangling ends and tails may perturb the duplex structure/solvation environment in some manner. }. This is likely due to 5' ends preferentially stacking on the core duplex compared to 3' ends. Furthermore, DNA nearest neighbor (NN) calculations that include dangling end contributions (Figure \ref{fig:AT-all_in_one}) predict that the specific nearest neighbors bonds in the 5' shifted state are more energetically favorable than those in the 3' shifted state \textcolor{purple}{Also, it may be worth stating that NN parameters agree with experimental melting thermodynamics for the studied sequences (cite Paul's JACS paper). The NN parameters are occasionally quite bad for some sequences.} \citep{Santalucia2004TM, SantaLucia1998AThermodynamics} (Note that these calculations are performed at higher salt concentration, and that salt seems to have a significant effect on the impact of inert tails on stability) \textcolor{purple}{You may want to be more specific about how counterions impact inert tail thermodynamics  (more destabilizing at low counterion) and that your simulations are at fairly high counterion concentration}. Taken together, these experimental results substantiate why we observe the single-shifted 5' state to be only slightly less stable relative to the aligned hybridized state. Furthermore, we expect implied timescales of the two leading modes to be on the same order of magnitude as these reflect the average lifetimes of these states. The stability of the single-shifted 3' state is notably lower than that of the single-shifted 5' state. In fact, the free energy is about equal to that of the double-shifted 5' state, despite it having less intact base pairs. This is partially explained by nearest neighbor calculations, however these do not account for the differential effects of the 5' and 3' inert tails or kinetic considerations which we will discuss more below \textcolor{purple}{another inaccuracy of DNA NN is that they assume an all-or-nothing hybridization, so if the melting of each of these isolated conformations contained varying deviation from all-or-nothing behavior, then there would be another factor unaccounted for. 

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/AT-all_in_one.PNG}
        \caption{a) Thermo value calculated from Santalucia data and from MSMs b) TICA projections of freen energies and state clustering c) and transitions probabilities between states \textcolor{purple}{Consider getting rid of enthalpy and entropy columns (can add to SI). Need to add State numbers next to names}}
        \label{fig:AT-all_in_one}
	\end{center}
\end{figure}

\subsubsection{\label{sec:Results}Transition probabilities between states}

Beyond state probabilities and free energy approximations, the macrostate MSM also yields discrete kinetic information in the form of transition probabilities between states. Figure \ref{fig:AT-all_in_one} shows the probability of moving from one state to another within the MSM lag time (1.2 ns). All transition probabilities are higher when moving towards a more aligned state than towards a more shifted state, suggesting that these metastable shifted states play a more significant role in facilitating the hybridization process than dissociation. Furthermore, we see equal or higher transition rates from shifted states to the dissociated state than to more aligned states, indicating that the slithering-hybridization process is frequently disrupted by complete dissociation \textcolor{purple}{This is consisting with the GCGCGC sequence of Wales, right? They saw a large barrier to association from the double shifted state}. In particular, we observe that the transition probability from double-shifted 3' state to the dissociated state is 6x higher than to the neighboring single-shifted 3' state. Although there is an approximately equal probability of transitioning from the dissociated state to the hybridized and 5' shifted states, there is a 2x lower probability of moving to the 3' shifted states. These asymmetric kinetic effects may have some role in the substantially lower state populations we see above. From these observations we conclude that shifted states are important in facilitating the hybridization process, but that increasingly shifted states (in particular when shifted in the 3' direction) are more likely to dissociate than to evolve into a fully aligned hybridized state \textcolor{purple}{Based on the thermo studies of dangling ends and inert tails, it seems like the role of shifted states may depend on salt conditions, this could be an interesting future direction.}.

>>>>

Over the last couple decades, DNA has proved to be much more than a vessel for genetic information. From sensing, to computing, to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen applications \citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems.pdf, Rothemund2006FoldingPatterns, Gu2010ALine}. Recently, structural DNA nanotechnology has enabled self-assembly on micro to milli scales, and dynamic DNA nanotechnology has been used to perform basic calculation and to probe single molecules via temporal DNA signatures\citep{MhatreV.HoJi-AnnLee2012NIHAccess, Bui2018, Shah2019}. \noter{Please correct reference ``MhatreV.HoJi-AnnLee2012NIHAccess''} Both technologies rely on the hybridization reaction between complementary DNA strands and leverage the flexibility of shorter DNA oligomers to participate in these reactions. Although many experimental and computational studies have rigorously explored DNA dynamical phenomena such as hybridization, hairpin formation, and single base pair flipping, the sequence-dependent mechanisms of hybridization and dissociation dynamics are not fully understood \citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition, Porschke1973ThermodynamicsPairs, Chen2007InfluenceHybridization, Craig1971ElaxationOligon}. Moreover, it is unclear the extent to which these processes evolve in an "all-or-nothing" fashion or if an ensemble meta-stable states facilitates the transition. Recent breakthrough studies have coupled experimental techniques with machine learning and MD simulations to investigate and predict sequence-dependent kinetics \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence}. Where these studies focus on association and dissociation kinetics alone, we broaden our analysis into higher order dynamical processes and meta-stable intermediates. The stability of certain intermediates states, such as out-of-register or  base pairing in repetitive sequences, has been well documented in previous computational studies \citep{Phys2014, Xiao2019, Maciejczyk2014DNAModel}) \textcolor{magenta}{No experimental has been done to hint at these dynamics (difficult to differentiate from fraying), but they've been discuss in many computational works and seem likely to occur}. Furthermore, frayed structures and dynamics have been investigated in numerous computational and experimental studies \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix}. Sandstead et al. highlighted the role of these dynamics during duplex dissociation, where the stability of frayed states was dictated by G:C bond placement in the 10-mer oligonucleotide sequence \citep{Sanstead2016}. In this work, we study the same four sequences explored by Sandstead et al in an effort to uncover the sequence-dependent dynamics and their relation to  metastable structures mentioned above.

\textcolor{magenta}{I would suggest re-organizing this paragraph to emphasize how the dynamic insight from various experiments is limited and therefore high-level simulations are required to understand DNA dynamics in detail. Then, you can summarize why simulating DNA hybridization is difficult and the previously employed techniques} Given the long timescales on which DNA hybridization and dissociation events occur, the study of these processes are generally not amenable to direct simulation techniques \citep{Phys2014}. Instead, many previous studies of DNA hybridization have employed accelerated sampling methods such as umbrella sampling \citep{Schmitt2013ExploringSurface} transition path sampling \citep{Sambriski2009}, and forward flux sampling \citep{Schmitt2013ExploringSurface, Sambriski2009,  Hoefert2011MolecularOligonucleotides, Romano2013DNADependence}. Other computational studies use dramatically elevated temperature or denaturing solvent concentrations induce one-way dissociation events \citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding}. Experimental studies often employ temperature-jump, salt-jump, pH-jump, or other perturbative methods to drive DNA out of equilibrium and monitor relaxation processes in one direction \citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971ElaxationOligon, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization}. Single molecule diffusion and tethered multifluorophore assays facilitate equilibrium analysis, however these present technical difficulties and are hampered by data collection rates \citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices}. Förster resonance energy transfer (FRET) analysis, particularly when coupled with methods mentioned above, provides additional resolution, but it unclear how fluorescent tags may interfere with the dynamics of short nucleotides \citep{Morrison1993SensitiveSolution}.

@@

We initialized four sequences previously investigated by Sandstead et al. -- 5'ATATATATAT3' (AT-all), 5'GATATATATC3' (GC-end), 5'ATATGCATAT3' (GC-core), and 5'ATGATATCAT3' (GC-mix) -- along with their complementary strands according to 3SPN.2 documentation \citep{Sanstead2016, Phys2014}. We initialized explicit ions such that 240 mM NaCl and 18 mM MgCl2 \textcolor{magenta}{So these are initially in the bulk and then may interact with the DNA? Are these the effective concentrations after you shrink the box?} were added to the box in addition to 18 Na counter ions to balance the charge from the 9 phosphate groups in each oligonucleotide backbone \citep{Hinckley2015}. In order to maximize concentration without allowing strands to see each other through periodic boundaries, we set the box size just larger than the sum of the maximum end-to-end extension length of a single strand and the force cutoff (using Ewald summation method is set at 20 A). This translated to a box size of 77.74 Å and an effective oligo concentration of 7 mM. We used an Ewald potential to calculate long range Coulombic interaction between DNA and ions. We maintained a Debye-Huckel screening potential with an ionic strength of 240 mM -- corresponding to NaCl concentration -- to account for phosphate backbone interactions. These electrostatics preserve the persistence length and intrinsic curvature of DNA while taking the effects of ion-DNA interactions\citep{Hinckley2015}. We run our simulations in the NVE ensemble and fix temperature via a Langevin thermostat in order to model implicit solvent interactions \citep{Schneider1978Molecular-dynamicsTransitions}. Simulation temperature was determined by sequence specific melting temperature such that strands were equally likely to spend time in the hybridized or dissociated state. Tmelt simulations were run for $3*10^{8}$ time steps, equating to 4.5$\mu$s simulation time, and frames were saved every 30 ps. A 15 fs timestep was used for all simulations runs. For each sequences, 100 simulations were performed in parallel, consuming about 32 serial cpu-hours of computation time for each independent simulation. An approximate Tmelt Boltzman distribution was replicated by initializing half of runs from the hybridized state and half from a random dissociated state. In order to allow for further equilibration, the first third (1.5 $\mu$s) of each simulation was removed  \textcolor{magenta}{Why 1.5 microseconds? Is this arbitrary or is there a good reason to remove 1.5?}, resulting in 100 x 100000 frames and a total of 300$\mu$s simulation time per sequence.

@@

All intermolecular pairwise distances for both oligonucleotides were calculated at each frame using the MDtraj software package \citep{McGibbon2015MDTraj:Trajectories}. Based on the self-complementary nature of each sequence, we averaged permutable distances (45 pairs in total) together. This permutation reduction follows a similar procedure used in TICAgg coordinates construction \citep{Sengupta2019AutomatedSelf-assembly}. The VAMP-2 scoring method \textcolor{magenta}{Since the VAMP-2 scoring comes up a lot, can you maybe another sentence or two on how exactly this works? Citing the Noe paper helps, but I think experimentalists may get a bit lost with it}, which represents the sum of squared estimates for the transfer operator, was employed to evaluate the quality of this feature set \citep{Mardt2018VAMPnetsKinetics}. We compared the VAMP-2 score for the permutation-free distances to the complete set of intermolecular distances and found only a small loss in kinetic variance when using the reduced data set. Furthermore, we found an increase in VAMP-2 score when using reciprocal pairwise distances and chose these reciprocal permutation-free coordinates as a consistent feature set. These features were normalized and passed into sequence-specific SRVs.

@@

\section{\label{sec:Results}Results}

\subsection{AT-all}

\subsubsection{SRV optimization and analysis}

In our analysis, we found that the AT-all sequence, given its repetitive structure and lack of GC-content, produced the cleanest dynamics and displayed a canonical spectral gap between modes. For this reason, we lead our discussion with this sequence and use it as a case study to work through our SRV-MSM pipeline step-by-step. Our first task was to identify the SRV lag time that was longer than the intrinsic Markov timescales of the system, yet short enough to resolve the dynamics of interest \citep{Phys2018MarkovValidation}. We found that most implied timescales (with the exception of the leading timescale) converge at a lag time of 600 ns. We kept a looser constraint on the convergence of the leading mode as it reflects the less Markovian dynamic of strand association and dissociation \textcolor{magenta}{Not sure how to describe this or if it warrants further discussion?} . It should be noted that our lag time selection was informed by the other sequences as well to maintain consistency. Next we selected an optimal number of SRV components to include in our analysis. After a certain point, higher order dynamical modes provide diminishing contributions the overall kinetic variance as measured by the Vamp-2 score, and the model can begin fitting on statistical noise in the trajectory data instead of the true dynamics \citep{McGibbon2015VariationalKinetics}. It is also more difficult to perform kmeans clustering on a high dimensional space, especially when those higher dimensions are less kinetically relevant \citep{Pande2010EverythingAsk}. For these reasons, the number of slow SRV components should be carefully selected based on the specific system of interest. As shown in (Figure \ref{fig:srv_crossval}), we see a clear convergence of the VAMP-2 score after five slow modes and select these modes as our optimized SRV basis. Next, we seek to interpret the physical relevance of these leading modes by plotting the Pearson correlation of each mode with the 100 intermolecular distances between strands. The quantitative meaning of these coordinates can be difficult to interpret given their nonlinear relationship to the SRV collective variables, but the relative different between these correlations shows which coordinates are most affected by (or effective on) each process. For example, the first slow mode shows a significant positive correlation to each distance and the strongest correlation with matching base pair distances (shown along the main diagonal). Given these relationships and the substantially longer timescale of this process, we can deduce that this leading mode corresponds to the dynamics of the overall hybridization and dissociation process. The next four SRV components all show a relatively high correlation along different offset diagonals. These diagonals correspond to the intermolecular distances between shifted base pairs and point to the existence of gradually faster shifting or "slithering" processes \textcolor{magenta}{I suppose the GCGCGC sequence in the Wales paper provides an interesting comparison to AT-all
}. Slow slithering mechanisms have similarly been reported in simulation studies to occur on orders of magnitude longer timescales than underlying fast dynamics \citep{Markegard2015, Xiao2019, Maciejczyk2014DNAModel}.

\subsubsection{SRV-MSM construction and optimization}

From this analysis we can determine that the slowest dynamics are fully characterized by the hybridization process and slithering behavior of out of register base pairs. Although this is qualitatively informative, we can access a more holistic picture of sequence kinetics and thermodynamics by using these SRV coordinates as a basis on which to construct an MSM. Because these coordinates are already capturing a majority of the system's kinetic variance, they serve as an ideal basis on which to group frames into microstates. We performed k-means clustering, and optimized the number of microstates at 200 by monitoring VAMP-2 score. Next, we selected an MSM lag time in a similar fashion to our SRV lag time selection process. In Figure \ref{fig:AT-all_dynamic} we compare the convergence of SRV-MSM and TICA-MSM timescales, where SRV-MSM timescales converge consistently faster and to higher values of all leading modes. This enables us to select a shorter lag time and build a higher resolution model than we could from an analogous TICA basis. Using an MSM lag time of 1.2 ns, we then built a Bayesian MSM to calculate transition probability matrix between each microstate. Finally, PCCA+ spectral clustering was implemented to group these microstates into macrostates that each represent a collection of metastable structures. Previous work have used a common set of microstates and/or performed manual clustering of microstates based on physical read outs from simulation data (stacking score, energies, etc) \citep{Pinamonti2017PredictingModels}, \citep{PinamontiAnalyzedModels}. Although these techniques are useful for performing comparisons between sequences, we saw better results when optimizing MSMs to capture the most detail of sequence individually and thus developed an independent set of microstates and macrostates for each sequence. For AT-all, we kept to the convention of clustering into n+1 macrostates, where n is the number of slow components captured by the MSMs. To visualize these six macrostates, we project the data into the two leading TICA coordinates. Although SRVs outperform these coordinates for the purpose of MSM construction, TICA represent good high variance collective variables on which to visualize free energies and state assignments \citep{Sidky} \textcolor{magenta}{Still not sure why you use tICA here, does not nonlinear nature make them harder to include physically? Should these be in SI}. It is clear in Figure \ref{fig:AT-all_in_one} that the PCCA state assignments are capturing free energy minima in TICA space as independent states. After assigning these macrostates we can then calculate their relative probabilities and free energies, visualize representative molecular renderings, and estimates transition probabilities between states. In this final step, we use a minimum flux cutoff of 2e-6 in order to mitigate erroneous quick transition or skipping between states.

@@

\subsubsection{Implied timescales for TICA-MSMs for }

In our coarse-grained SRV-MSM, we observe an "aligned hybridized" state, dissociated state, and four "shifted" states characterized by different combinations of out-of-register base pairings. These shifted states consist of 2 or 4 base pair shifting (single or doubled shifted) in either the 5' or 3' direction with varied stability and transition fluxes between kinetically neighboring states. This state decomposition is expected given the repetitive nature of the AT-all sequence and previous computational results that find shifted conformations form "deep kinetic traps" \citep{Xiao2019}, \citep{Phys2014}. We found a substantial difference in thermodynamic stability and kinetic behavior between the 5'A shifted states and 3'T shifted states. This difference can be accounted for by examining experimental studies on the thermodynamics of "dangling ends" -- unpaired bases adjacent to the paired duplex -- and "inert tails" -- free bases that extend beyond the dangling end \citep{Michele2014EHybridization}. Dangling 5' ends have a consistently stabilizing effect, and inert tails decrease stability as they increase in length \textcolor{magenta}{The effect of inert tails seems sequence dependent and not always destabilizing (Benight), especially at high ionic strength}. It has been shown that two base pair 5' dangling ends (or dangling ends with one inert tail) have higher melting temperature and are enthalpically favorable compared to 3' ends \citep{Senior1988InfluenceDuplexes, Dickman2012ThermodynamicDNAs} \textcolor{magenta}{The Benight paper also shows that 3'-dangling ends and tails alter the change in heat capacity (dCp) upon hybridization much more than 5'. Additionally, their analysis suggests 3'-dangling ends lead to a reduction in bound counterions relative to 5'. Taken together, it would seem that 3'-dangling ends and tails may perturb the duplex structure/solvation environment in some manner. }. This is likely due to 5' ends preferentially stacking on the core duplex compared to 3' ends. Furthermore, DNA nearest neighbor (NN) calculations that include dangling end contributions (Figure \ref{fig:AT-all_in_one}) predict that the specific nearest neighbors bonds in the 5' shifted state are more energetically favorable than those in the 3' shifted state \textcolor{magenta}{Also, it may be worth stating that NN parameters agree with experimental melting thermodynamics for the studied sequences (cite Paul's JACS paper). The NN parameters are occasionally quite bad for some sequences.} \citep{Santalucia2004TM, SantaLucia1998AThermodynamics} (Note that these calculations are performed at higher salt concentration, and that salt seems to have a significant effect on the impact of inert tails on stability) \textcolor{magenta}{You may want to be more specific about how counterions impact inert tail thermodynamics  (more destabilizing at low counterion) and that your simulations are at fairly high counterion concentration}. Taken together, these experimental results substantiate why we observe the single-shifted 5' state to be only slightly less stable relative to the aligned hybridized state. Furthermore, we expect implied timescales of the two leading modes to be on the same order of magnitude as these reflect the average lifetimes of these states. The stability of the single-shifted 3' state is notably lower than that of the single-shifted 5' state. In fact, the free energy is about equal to that of the double-shifted 5' state, despite it having less intact base pairs. This is partially explained by nearest neighbor calculations, however these do not account for the differential effects of the 5' and 3' inert tails or kinetic considerations which we will discuss more below \textcolor{magenta}{another inaccuracy of DNA NN is that they assume an all-or-nothing hybridization, so if the melting of each of these isolated conformations contained varying deviation from all-or-nothing behavior, then there would be another factor unaccounted for. 

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/AT-all_in_one.PNG}
        \caption{a) Thermo value calculated from Santalucia data and from MSMs b) TICA projections of freen energies and state clustering c) and transitions probabilities between states \textcolor{magenta}{Consider getting rid of enthalpy and entropy columns (can add to SI). Need to add State numbers next to names}}
        \label{fig:AT-all_in_one}
	\end{center}
\end{figure}

\subsubsection{Transition probabilities between states}

Beyond state probabilities and free energy approximations, the macrostate MSM also yields discrete kinetic information in the form of transition probabilities between states. Figure \ref{fig:AT-all_in_one} shows the probability of moving from one state to another within the MSM lag time (1.2 ns). All transition probabilities are higher when moving towards a more aligned state than towards a more shifted state, suggesting that these metastable shifted states play a more significant role in facilitating the hybridization process than dissociation. Furthermore, we see equal or higher transition rates from shifted states to the dissociated state than to more aligned states, indicating that the slithering-hybridization process is frequently disrupted by complete dissociation \textcolor{magenta}{This is consisting with the GCGCGC sequence of Wales, right? They saw a large barrier to association from the double shifted state}. In particular, we observe that the transition probability from double-shifted 3' state to the dissociated state is 6x higher than to the neighboring single-shifted 3' state. Although there is an approximately equal probability of transitioning from the dissociated state to the hybridized and 5' shifted states, there is a 2x lower probability of moving to the 3' shifted states. These asymmetric kinetic effects may have some role in the substantially lower state populations we see above. From these observations we conclude that shifted states are important in facilitating the hybridization process, but that increasingly shifted states (in particular when shifted in the 3' direction) are more likely to dissociate than to evolve into a fully aligned hybridized state \textcolor{magenta}{Based on the thermo studies of dangling ends and inert tails, it seems like the role of shifted states may depend on salt conditions, this could be an interesting future direction.}.
